---
title: "Cell-autonomous functional inference of Cluster 1 and Cluster 2 Astrocyte Gene Signatures"
author: "Arpy"
date: '2024-09-24'
output: html_document
---

```{r libraries and functions, message=FALSE}

library(Seurat)
library(tidyverse)
library(readxl)
library(dplyr)
library(clusterProfiler)
library(ReactomePA)
library(enrichplot)
library(ggtree)
library(scales)

source("~/OHSU Dropbox/Saunders Lab's shared workspace/arpy/manuscripts/2023_Thai2P4M_FeigeYoung/ms_analyses/1_karl_analysis/r_functions_paths/_color_palettes_tha2P4M.R")
source("~/OHSU Dropbox/Saunders Lab's shared workspace/arpy/manuscripts/2023_Thai2P4M_FeigeYoung/ms_analyses/1_karl_analysis/r_functions_paths/_karl_tha2P4M_functions.R")
source("~/OHSU Dropbox/Saunders Lab's shared workspace/arpy/manuscripts/2023_Thai2P4M_FeigeYoung/ms_analyses/1_karl_analysis/r_functions_paths/_karl_tha2P4M_filepaths.R")
```

#0. Load Data and Define Save Paths
```{r load data, echo=FALSE, attr.warning= FALSE}

###Load
#DE Meta1 results 
de.meta.path <- "~/OHSU Dropbox/Saunders Lab's shared workspace/arpy/manuscripts/2023_Thai2P4M_FeigeYoung/ms_analyses/1_karl_analysis/Data/de/de_with_genemetadata/"

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta1            <- readRDS(paste0(de.meta.path, "DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta1.RDS"))
DE.tha05.vs.NI.CellType                                     <- readRDS(DE.tha05.vs.NI.CellType.path)



#Tha Seurat object and infections.summary
tha.celltype.infections.summary <- read_rds(tha.celltype.infections.summary.path)
tha.seurat                      <- read_rds(tha.seurat.path)



#DE Astrocyte Data from Published Sources
prakash.path <- "~/OHSU Dropbox/Saunders Lab's shared workspace/arpy/manuscripts/2023_Thai2P4M_FeigeYoung/ms_analyses/1_karl_analysis/Data/liddelow/2023_prakash/"
hasel.path   <- "~/OHSU Dropbox/Saunders Lab's shared workspace/arpy/manuscripts/2023_Thai2P4M_FeigeYoung/ms_analyses/1_karl_analysis/Data/liddelow/2021_hasel/"

#table 5 describes a comparison of the rodent (rat) interferon-responsive reactive astrocyte (IRRA) proteome and transcriptome
prakash_table5.concordant    <- readxl::read_excel(path = paste0(prakash.path, "glia24494-sup-0011-tables5.xlsx"), sheet = 1)
prakash_table5.discordant    <- readxl::read_excel(path = paste0(prakash.path, "glia24494-sup-0011-tables5.xlsx"), sheet = 2)
prakash_table5.protein_only  <- readxl::read_excel(path = paste0(prakash.path, "glia24494-sup-0011-tables5.xlsx"), sheet = 3)
prakash_table5.rna_only      <- readxl::read_excel(path = paste0(prakash.path, "glia24494-sup-0011-tables5.xlsx"), sheet = 4)
prakash_table5.neither       <- readxl::read_excel(path = paste0(prakash.path, "glia24494-sup-0011-tables5.xlsx"), sheet = 5)

#table 3 describes the proteome differences in human astrocytes triggered with human IL1α, TNF and human C1Q
prakash_table3               <- readxl::read_excel(path = paste0(prakash.path, "glia24494-sup-0009-tables3.xls"), sheet = 1)


hasel.cluster_enriched       <- readxl::read_excel(path = paste0(hasel.path, "41593_2021_905_MOESM6_ESM.xlsx"), sheet = 1)

```
```{r define output paths, echo = FALSE}
de.data.path    <- "~/OHSU Dropbox/Saunders Lab's shared workspace/arpy/manuscripts/2023_Thai2P4M_FeigeYoung/ms_analyses/1_karl_analysis/Data/de/"
astro.gsea.path <- "~/OHSU Dropbox/Saunders Lab's shared workspace/arpy/manuscripts/2023_Thai2P4M_FeigeYoung/ms_analyses/1_karl_analysis/Plots/6_astrocytes/gsea/"
```

#1a. PreProcess and Save Published DE Datasets
```{r preprocess published DE datasets, echo=FALSE}
#Prakash Table 3 Preprocessing
colnames(prakash_table3) <- prakash_table3[1,]
prakash_table3           <- prakash_table3[-c(1:2),-1]

prakash.3.keeper.cols <- c("p value", "TICI_VEH fold change, log 2", "Gene names")

prakash_table3 <-
  prakash_table3 %>%
  select(prakash.3.keeper.cols) %>%
  mutate(Significance_Prakash_3_HumanProteome = case_when(`p value` < 0.05 ~ "Significant", 
                                                          `p value` == "4.7907535209631978e-05" ~ "Significant",
                                                          TRUE ~ "NotSignificant")) %>%
  rename(gene = `Gene names`)


#Prakash Table 5 Preprocessing
colnames(prakash_table5.concordant) <- prakash_table5.concordant[1,]
prakash_table5.concordant <- prakash_table5.concordant[-1,-1]

colnames(prakash_table5.discordant) <- prakash_table5.discordant[1,]
prakash_table5.discordant <- prakash_table5.discordant[-1,-1]

colnames(prakash_table5.protein_only) <- prakash_table5.protein_only[1,]
prakash_table5.protein_only <- prakash_table5.protein_only[-1,-1]

colnames(prakash_table5.rna_only) <- prakash_table5.rna_only[1,]
prakash_table5.rna_only <- prakash_table5.rna_only[-1,-1]

colnames(prakash_table5.neither) <- prakash_table5.neither[1,]
prakash_table5.neither <- prakash_table5.neither[-1,-1]      

prakash.5.keeper.cols <- c("gene_names", "log2FC", "pval", "log2FoldChange", "padj", "Significance")

prakash_table5.concordant.simple <-
  prakash_table5.concordant %>%
  select(prakash.5.keeper.cols)

prakash_table5.discordant.simple <-
  prakash_table5.discordant %>%
  select(prakash.5.keeper.cols)

prakash_table5.protein_only.simple <-
  prakash_table5.protein_only %>%
  select(prakash.5.keeper.cols)

prakash_table5.rna_only.simple <-
  prakash_table5.rna_only %>%
  select(prakash.5.keeper.cols)

prakash_table5.neither.simple <-
  prakash_table5.neither %>%
  select(prakash.5.keeper.cols)

prakash_table5.ALL.simple <- rbind(prakash_table5.concordant.simple, prakash_table5.discordant.simple, prakash_table5.protein_only.simple, prakash_table5.rna_only.simple, prakash_table5.neither.simple)

prakash_table5.ALL.simple <- 
  prakash_table5.ALL.simple %>%
  mutate(gene_mouse = gene_names) %>%
  mutate(gene = toupper(gene_names)) %>%
  relocate(gene) %>%
  select(-gene_names) %>%
  mutate(Significance_Prakash_5_RodentProteome_vs_Transcriptome = case_when(Significance == "Both (concordant)(n = 67)" ~ "Concordant",
                                  Significance == "Protein only (n = 14)" ~ "Protein_Only",
                                  Significance == "RNA only (n = 966)" ~ "RNA_Only",
                                  Significance == "Neither (n = 1099)" ~ "Neither",
                                  TRUE ~ "Not_Included"))


#Hasel DE Clusters Preprocessing
hasel.cluster_enriched <-
  hasel.cluster_enriched %>%
  mutate(gene_mouse = Gene_name) %>%
  mutate(gene = toupper(Gene_name)) %>%
  relocate(gene) %>%
  select(-Gene_name)

hasel.cluster_enriched.cluster8 <-
  hasel.cluster_enriched %>%
  filter(Cluster == 8)

```


#1b. Push gene metadata to the DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta1
```{r gene score scatter plots, echo=FALSE}

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta1 %>%
  mutate(Hasel_Cluster8 = case_when(gene %in% hasel.cluster_enriched.cluster8$gene ~ "Cluster_8", TRUE ~ "NotCluster_8"))

prakash_table5.ALL.simple_temp <-
  prakash_table5.ALL.simple %>%
  select(gene, Significance_Prakash_5_RodentProteome_vs_Transcriptome)

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 <- left_join(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, prakash_table5.ALL.simple_temp, by = "gene")

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 %>%
  mutate(Significance_Prakash_5_RodentProteome_vs_Transcriptome=replace_na(Significance_Prakash_5_RodentProteome_vs_Transcriptome, "Not_Included"))

prakash_table3_temp <-
  prakash_table3 %>%
  select(gene, Significance_Prakash_3_HumanProteome)

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 <- left_join(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, prakash_table3_temp, by = "gene")

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 %>%
  mutate(Significance_Prakash_3_HumanProteome=replace_na(Significance_Prakash_3_HumanProteome, "Not_Included"))

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 %>%
  mutate("Astro_0_vs_2_Sig" = case_when(p_val_adj < 0.01 & avg_logFC < 0 ~ "Astro_0_UP", 
                                        p_val_adj < 0.01 & avg_logFC > 0 ~ "Astro_2_UP",
                                        TRUE ~ "NotSignificant"))

#Save
de.meta.path <- "~/OHSU Dropbox/Saunders Lab's shared workspace/arpy/manuscripts/2023_Thai2P4M_FeigeYoung/ms_analyses/1_karl_analysis/Data/de/de_with_genemetadata/"
saveRDS(object = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, file = paste0(de.meta.path, "DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2.RDS"))


```

#2a. Volcano Plotting on Published DE Datasets 
```{r volcano plots with metadata, echo=TRUE}
#Load
de.meta.path <- "~/OHSU Dropbox/Saunders Lab's shared workspace/arpy/manuscripts/2023_Thai2P4M_FeigeYoung/ms_analyses/1_karl_analysis/Data/de/de_with_genemetadata/"
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 <- readRDS(file = paste0(de.meta.path, "DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2.RDS"))

#save path
astro_de.path <- "~/OHSU Dropbox/Saunders Lab's shared workspace/arpy/manuscripts/2023_Thai2P4M_FeigeYoung/ms_analyses/1_karl_analysis/Plots/6_astrocytes/volcano/"

###Hasel et al Cluster 8
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 %>%
  arrange(match(Hasel_Cluster8, c("NotCluster_8", "Cluster_8")))

Volcano_Hasel_Cluster8_reference <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, aes(x = avg_logFC, y = -log10(p_val_adj), color = Astro_0_vs_2_Sig, fill = Hasel_Cluster8, alpha = Hasel_Cluster8, size = Hasel_Cluster8))+
     geom_point(shape = 21)+
     scale_color_manual(values = c("red", "red", "grey"))+
     scale_fill_manual(values = c("#2b8cbe", "grey"))+
     scale_size_manual(values = c(3, 1))+
     scale_alpha_manual(values = c(0.9,0.15))+
     #scale_size_manual(values = c(1.5,1.5,0.5,0.5))+
     #guides(color = "none", fill = "none", alpha = "none", size = "none")+
     geom_hline(yintercept=2, linetype = 2)+
     scale_x_continuous(limits = c(-5.5, 5.5))+
     scale_y_continuous(limits = c(0, 25))+
     theme_bw()+
      theme(axis.title=element_blank(),
        strip.background=element_blank())

Volcano_Hasel_Cluster8 <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, aes(x = avg_logFC, y = -log10(p_val_adj), color = Astro_0_vs_2_Sig, fill = Hasel_Cluster8, alpha = Hasel_Cluster8, size = Hasel_Cluster8))+
     geom_point(shape = 21)+
     scale_color_manual(values = c("red", "red", "grey"))+
     scale_fill_manual(values = c("#2b8cbe", "grey"))+
     scale_size_manual(values = c(3, 1))+
     scale_alpha_manual(values = c(0.9,0.15))+
     guides(color = "none", fill = "none", alpha = "none", size = "none")+
     geom_hline(yintercept=2, linetype = 2)+
     scale_x_continuous(limits = c(-5.5, 5.5))+
     scale_y_continuous(limits = c(0, 25))+
     theme_bw()+
      theme(axis.title=element_blank(),
        strip.background=element_blank())

#Save
ggsave(paste0(astro_de.path, "Volcano_Hasel_Cluster8.png"),            Volcano_Hasel_Cluster8,           device='png', width = 3, height=2.25, units='in', dpi=600)
ggsave(paste0(astro_de.path, "Volcano_Hasel_Cluster8_reference.png"),  Volcano_Hasel_Cluster8_reference, device='png', width = 30, height=22.5, units='in', dpi=600)


###Prakesh et al Table 5 describes a comparison of the rodent (rat) interferon-responsive reactive astrocyte (IRRA) proteome and transcriptome
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 %>%
  arrange(match(Significance_Prakash_5_RodentProteome_vs_Transcriptome, c("Not_Included", "Neither", "RNA_Only", "Protein_Only", "Concordant")))

Volcano_Prakesh_Table5_reference <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, aes(x = avg_logFC, y = -log10(p_val_adj), color = Astro_0_vs_2_Sig, fill = Significance_Prakash_5_RodentProteome_vs_Transcriptome, alpha = Significance_Prakash_5_RodentProteome_vs_Transcriptome, size = Significance_Prakash_5_RodentProteome_vs_Transcriptome))+
     geom_point(shape = 21)+
     scale_color_manual(values = c("red", "red", "grey"))+
     scale_fill_manual(values = c("#2b8cbe", "grey", "grey", "#d95f02", "#7570b3"))+
     scale_size_manual(values = c(3, 1, 1, 3, 3))+
     scale_alpha_manual(values = c(0.9,0.15, 0.15, 0.9, 0.9))+
     geom_hline(yintercept=2, linetype = 2)+
     scale_x_continuous(limits = c(-5.5, 5.5))+
     scale_y_continuous(limits = c(0, 200))+
     theme_bw()+
      theme(axis.title=element_blank(),
        strip.background=element_blank())

Volcano_Prakesh_Table5 <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, aes(x = avg_logFC, y = -log10(p_val_adj), color = Astro_0_vs_2_Sig, fill = Significance_Prakash_5_RodentProteome_vs_Transcriptome, alpha = Significance_Prakash_5_RodentProteome_vs_Transcriptome, size = Significance_Prakash_5_RodentProteome_vs_Transcriptome))+
     geom_point(shape = 21)+
     scale_color_manual(values = c("red", "red", "grey"))+
     scale_fill_manual(values = c("#2b8cbe", "grey", "grey", "#d95f02", "#7570b3"))+
     scale_size_manual(values = c(2, 1, 1, 2, 2))+
     #scale_alpha_manual(values = c(0.9,0.15, 0.15, 0.9, 0.9))+
     scale_alpha_manual(values = c(1,0.15, 0.15, 0.3, 0.3))+
     guides(color = "none", fill = "none", alpha = "none", size = "none")+
     geom_hline(yintercept=2, linetype = 2)+
     scale_x_continuous(limits = c(-5.5, 5.5))+
     scale_y_continuous(limits = c(0, 100))+
     theme_bw()+
      theme(axis.title=element_blank(),
        strip.background=element_blank())

Volcano_Prakesh_Table5_Concordant_reference <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, aes(x = avg_logFC, y = -log10(p_val_adj), color = Astro_0_vs_2_Sig, fill = Significance_Prakash_5_RodentProteome_vs_Transcriptome, alpha = Significance_Prakash_5_RodentProteome_vs_Transcriptome, size = Significance_Prakash_5_RodentProteome_vs_Transcriptome))+
     geom_point(shape = 21)+
     scale_color_manual(values = c("red", "red", "grey"))+
     scale_fill_manual(values = c("#2b8cbe", "grey", "grey", "grey", "grey"))+
     scale_size_manual(values = c(3, 1, 1, 1, 1))+
     #scale_alpha_manual(values = c(0.9,0.15, 0.15, 0.9, 0.9))+
     #scale_alpha_manual(values = c(1,0.15, 0.15, 0.15, 0.15))+
     guides(color = "none", fill = "none", alpha = "none", size = "none")+
     geom_hline(yintercept=2, linetype = 2)+
     scale_x_continuous(limits = c(-5.5, 5.5))+
     scale_y_continuous(limits = c(0, 25))+
     theme_bw()+
      theme(axis.title=element_blank(),
        strip.background=element_blank())


Volcano_Prakesh_Table5_Concordant <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, aes(x = avg_logFC, y = -log10(p_val_adj), color = Astro_0_vs_2_Sig, fill = Significance_Prakash_5_RodentProteome_vs_Transcriptome, alpha = Significance_Prakash_5_RodentProteome_vs_Transcriptome, size = Significance_Prakash_5_RodentProteome_vs_Transcriptome))+
     geom_point(shape = 21)+
     scale_color_manual(values = c("red", "red", "grey"))+
     scale_fill_manual(values = c("#2b8cbe", "grey", "grey", "grey", "grey"))+
     scale_size_manual(values = c(3, 1, 1, 1, 1))+
     #scale_alpha_manual(values = c(0.9,0.15, 0.15, 0.9, 0.9))+
     scale_alpha_manual(values = c(1,0.15, 0.15, 0.15, 0.15))+
     guides(color = "none", fill = "none", alpha = "none", size = "none")+
     geom_hline(yintercept=2, linetype = 2)+
     scale_x_continuous(limits = c(-5.5, 5.5))+
     scale_y_continuous(limits = c(0, 25))+
     theme_bw()+
      theme(axis.title=element_blank(),
        strip.background=element_blank())

#Save
ggsave(paste0(astro_de.path, "Volcano_Prakesh_Table5.png"),                       Volcano_Prakesh_Table5,                                 device='png', width = 3,  height=2.25, units='in', dpi=600)
ggsave(paste0(astro_de.path, "Volcano_Prakesh_Table5_reference.png"),             Volcano_Prakesh_Table5_reference,                       device='png', width = 30, height=22.5, units='in', dpi=600)
ggsave(paste0(astro_de.path, "Volcano_Prakesh_Table5_Concordant.png"),            Volcano_Prakesh_Table5_Concordant,                      device='png', width = 3,  height=2.25, units='in', dpi=600)
ggsave(paste0(astro_de.path, "Volcano_Prakesh_Table5_Concordant_reference.png"),  Volcano_Prakesh_Table5_Concordant_reference,            device='png', width = 30, height=22.5, units='in', dpi=600)




###Prakesh et al Table 3 describes a human astrocytes interferon-responsive reactive astrocyte (IRRA) proteome
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 %>%
  arrange(match(Significance_Prakash_3_HumanProteome, c("Not_Included", "NotSignificant", "Significant")))

Volcano_Prakesh_Table3_reference <-
   ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, aes(x = avg_logFC, y = -log10(p_val_adj), color = Astro_0_vs_2_Sig, fill = Significance_Prakash_3_HumanProteome, alpha = Significance_Prakash_3_HumanProteome, size = Significance_Prakash_3_HumanProteome))+
     geom_point(shape = 21)+
     scale_color_manual(values = c("red", "red", "grey"))+
     scale_fill_manual(values = c("grey", "grey", "#2b8cbe"))+
     scale_size_manual(values = c(1, 1, 3))+
     scale_alpha_manual(values = c(0.15, 0.15, 1))+
     #guides(color = "none", fill = "none", alpha = "none", size = "none")+
     geom_hline(yintercept=2, linetype = 2)+
     scale_x_continuous(limits = c(-5.5, 5.5))+
     scale_y_continuous(limits = c(0, 25))+
     theme_bw()+
      theme(axis.title=element_blank(),
        strip.background=element_blank())

Volcano_Prakesh_Table3 <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, aes(x = avg_logFC, y = -log10(p_val_adj), color = Astro_0_vs_2_Sig, fill = Significance_Prakash_3_HumanProteome, alpha = Significance_Prakash_3_HumanProteome, size = Significance_Prakash_3_HumanProteome))+
     geom_point(shape = 21)+
     scale_color_manual(values = c("red", "red", "grey"))+
     scale_fill_manual(values = c("grey", "grey", "#2b8cbe"))+
     scale_size_manual(values = c(1, 1, 3))+
     scale_alpha_manual(values = c(0.15, 0.15, 1))+
     guides(color = "none", fill = "none", alpha = "none", size = "none")+
     geom_hline(yintercept=2, linetype = 2)+
     scale_x_continuous(limits = c(-5.5, 5.5))+
     scale_y_continuous(limits = c(0, 25))+
     theme_bw()+
      theme(axis.title=element_blank(),
        strip.background=element_blank())


#Save
ggsave(paste0(astro_de.path, "Volcano_Prakesh_Table3.png"),                       Volcano_Prakesh_Table3,                                 device='png', width = 3,  height=2.25, units='in', dpi=600)
ggsave(paste0(astro_de.path, "Volcano_Prakesh_Table3_reference.png"),             Volcano_Prakesh_Table3_reference,                       device='png', width = 30, height=22.5, units='in', dpi=600)
```
#3a. GSEA Analysis
```{r GSEA Analysis and plotting, echo = F}

###DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2
genes_to_keep <- tha.seurat@assays$RNA@meta.features %>%
  filter(vst.mean > 0.05) %>% rownames()

temp <- filter(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, gene %in% genes_to_keep,
  cell_type == 'Astrocyte') %>%
  mutate(stat = -log10(p_val_adj) * avg_logFC) %>%
  filter(!is.na(stat)) %>%
  arrange(desc(stat))

AstroLow_Cluster0.vs.AstroHigh_Cluster2        <- temp$stat
names(AstroLow_Cluster0.vs.AstroHigh_Cluster2) <- temp$gene

GSEA.AstroLow_Cluster0.vs.AstroHigh_Cluster2 <- clusterProfiler::gseGO(
  AstroLow_Cluster0.vs.AstroHigh_Cluster2,
  ont = 'BP',
  OrgDb = 'org.Hs.eg.db',
  keyType = 'SYMBOL',
  eps=1e-300
)

#Save
saveRDS(object = GSEA.AstroLow_Cluster0.vs.AstroHigh_Cluster2, file = paste0(de.data.path, "gsea/GSEA.AstroLow_Cluster0.vs.AstroHigh_Cluster2.RDS"))



###DE.tha2p4m.high.vs.low.Astrocyte
genes_to_keep <- tha.seurat@assays$RNA@meta.features %>%
  filter(vst.mean > 0.05) %>% rownames()

temp <- filter(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, gene %in% genes_to_keep,
  cell_type == 'Astrocyte') %>%
  mutate(stat = -log10(p_val_adj) * avg_logFC) %>%
  filter(!is.na(stat)) %>%
  arrange(desc(stat))

tha2p4m.high.vs.low.Astrocyte        <- temp$stat
names(tha2p4m.high.vs.low.Astrocyte) <- temp$gene

GSEA.tha2p4m.high.vs.low.Astrocyte <- clusterProfiler::gseGO(
  tha2p4m.high.vs.low.Astrocyte,
  ont = 'BP',
  OrgDb = 'org.Hs.eg.db',
  keyType = 'SYMBOL',
  eps=1e-300
)

#Save
saveRDS(object = GSEA.tha2p4m.high.vs.low.Astrocyte, file = paste0(de.data.path, "gsea/GSEA.tha2p4m.high.vs.low.Astrocyte.RDS"))



###DE.tha2P4M.AstroLow_Cluster0.vs.AstroHigh_Cluster2
genes_to_keep <- tha.seurat@assays$RNA@meta.features %>%
  filter(vst.mean > 0.05) %>% rownames()

temp <- filter(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2, gene %in% genes_to_keep,
  cell_type == 'Astrocyte') %>%
  mutate(stat = -log10(p_val_adj) * avg_logFC) %>%
  filter(!is.na(stat)) %>%
  arrange(desc(stat))

tha2P4M.AstroLow_Cluster0.vs.AstroHigh_Cluster2        <- temp$stat
names(tha2P4M.AstroLow_Cluster0.vs.AstroHigh_Cluster2) <- temp$gene

GSEA.tha2P4M.AstroLow_Cluster0.vs.AstroHigh_Cluster2 <- clusterProfiler::gseGO(
  tha2P4M.AstroLow_Cluster0.vs.AstroHigh_Cluster2,
  ont = 'BP',
  OrgDb = 'org.Hs.eg.db',
  keyType = 'SYMBOL',
  eps=1e-300
)

#Save
saveRDS(object = GSEA.tha2P4M.AstroLow_Cluster0.vs.AstroHigh_Cluster2, file = paste0(de.data.path, "gsea/GSEA.tha2P4M.AstroLow_Cluster0.vs.AstroHigh_Cluster2.RDS"))



#Visualize GSEA Results
#Simplify
GSEA.AstroLow_Cluster0.vs.AstroHigh_Cluster2_simplify       <- clusterProfiler::simplify(GSEA.AstroLow_Cluster0.vs.AstroHigh_Cluster2, cutoff=0.7, by="p.adjust", select_fun=min)

#Generate Plots
GSEA.tha2P4M.AstroLow_Cluster0.vs.AstroHigh_Cluster2.plot                         <- enrichplot::gseaplot2(GSEA.AstroLow_Cluster0.vs.AstroHigh_Cluster2, geneSetID = 1:10, ES_geom = "line")
GSEA.AstroLow_Cluster0.vs.AstroHigh_Cluster2_simplify.plot                        <- enrichplot::gseaplot2(GSEA.AstroLow_Cluster0.vs.AstroHigh_Cluster2_simplify, geneSetID = 1:10, ES_geom = "line")

```

###REACTOME Analysis

#4a. REACTOME Pre-Processing
```{r REACTOME analysis and plotting, echo=FALSE}
####################################
###REACTOME Analysis
###################################

#Preprocess DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2 for REACTOME analysis
#Add ENTREZID
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2.ENTREZID     <- bitr(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2$gene, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2.ENTREZID_tbl <- as_tibble(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2.ENTREZID)

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2 <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta2 %>%
  dplyr::rename(SYMBOL = gene)

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All <- left_join(x = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2.ENTREZID_tbl, y = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2, by = "SYMBOL")

#Filter UP and DOWN signficant genes
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All %>%
  dplyr::select(ENTREZID, p_val_adj, avg_logFC) %>%
  mutate(score = -log10(p_val_adj) * avg_logFC) %>%
  filter(p_val_adj < 0.01)

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean %>%
  filter(avg_logFC > 0)

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean %>%
  filter(avg_logFC < 0)

#generate gene-named LogFC vectors of significantly UP or DOWN regulated genes (UP = enriched in Cluster 2; DOWN = enriched in Cluster 1)
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_simple_UP           <- DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP$avg_logFC
names(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_simple_UP)    <- DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP$ENTREZID

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_simple_DOWN           <- DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN$avg_logFC
names(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_simple_DOWN)    <- DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN$ENTREZID


REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP   <- ReactomePA::enrichPathway(gene = names(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_simple_UP),     pvalueCutoff = 0.01, readable=TRUE)
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN <- ReactomePA::enrichPathway(gene = names(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_simple_DOWN),   pvalueCutoff = 0.01, readable=TRUE)

#Save Reactome Results
saveRDS(object = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP,   file = paste0(de.data.path, "gsea/REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP.RDS"))
saveRDS(object = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN, file = paste0(de.data.path, "gsea/REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN.RDS"))
```

#4b. REACTOME Pathway Plots
```{r Plot REACTOME Pathways, echo=FALSE}
#Visualize
#generate a logFC vector for all genes with ENTREZID names for visualzing key REACTOME Pathways
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_avg_logFC        <- DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All$avg_logFC
names(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_avg_logFC) <- DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All$ENTREZID

#Saving


# #set ggrepel() options
# options(ggrepel.max.overlaps = Inf)
# 
# ###UP
# #1. TollLikeReceptor3Cascade
# TollLikeReceptor3Cascade.pathway.plot <- ReactomePA::viewPathway(pathName = "Toll Like Receptor 3 (TLR3) Cascade", organism = "human", readable = TRUE, foldChange = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_avg_logFC)
# TollLikeReceptor3Cascade.pathway.plot <- TollLikeReceptor3Cascade.pathway.plot + 
#                                          scale_colour_gradient2(name = "fold change", low = "#006837", mid = "#e6f598", high = "#d73027")
# 
# 
# ggsave(paste0(astro.gsea.path, "TollLikeReceptor3Cascade.pathway.plot.png"), plot = TollLikeReceptor3Cascade.pathway.plot, device='png', width = 6, height=6, units='in', dpi=600)
# 
# 
# 
# #2. Toll-like Receptor Cascades
# TollLikeReceptorCascades.pathway.plot <- ReactomePA::viewPathway(pathName = "Toll-like Receptor Cascades", organism = "human", readable = TRUE, foldChange = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_avg_logFC)
# TollLikeReceptorCascades.pathway.plot <- TollLikeReceptorCascades.pathway.plot + 
#                                          scale_colour_gradient2(name = "fold change", low = "#006837", mid = "#e6f598", high = "#d73027")
# 
# 
# ggsave(paste0(astro.gsea.path, "TollLikeReceptorCascades.pathway.plot.png"), plot = TollLikeReceptorCascades.pathway.plot, device='png', width = 6, height=6, units='in', dpi=600)
# 
# 
# #3. tRNA processing
# tRNAProcessing.pathway.plot <- ReactomePA::viewPathway(pathName = "tRNA processing", organism = "human", readable = TRUE, foldChange = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_avg_logFC)
# tRNAProcessing.pathway.plot <- tRNAProcessing.pathway.plot + 
#                                          scale_colour_gradient2(name = "fold change", low = "#006837", mid = "#e6f598", high = "#d73027")
# 
# 
# ggsave(paste0(astro.gsea.path, "tRNAProcessing.pathway.plot.png"), plot = tRNAProcessing.pathway.plot, device='png', width = 6, height=6, units='in', dpi=600)
# 
# 
# 
# TollLikeReceptor3Cascade.pathway.plot <- ReactomePA::viewPathway(pathName = "Toll Like Receptor 3 (TLR3) Cascade", organism = "human", readable = TRUE, foldChange = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP)
# TollLikeReceptor3Cascade.pathway.plot <- TollLikeReceptor3Cascade.pathway.plot + 
#                                          scale_colour_gradient2(name = "fold change", low = "#006837", mid = "#e6f598", high = "#d73027")
# 
# #DOWN
# SignalingbyROBOreceptors.pathway.plot <- ReactomePA::viewPathway(pathName = "Signaling by ROBO receptors", organism = "human", readable = TRUE, foldChange = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_simple)
# SignalingbyROBOreceptors.pathway.plot <- SignalingbyROBOreceptors.pathway.plot + 
#                                          scale_colour_gradient2(name = "fold change", low = "#006837", mid = "#e6f598", high = "#d73027")
# 

```

#4c. REACTOME Cluster/Tree Plots
```{r Cluster and Plot REACTOME analyses based on similar gene sets using Tree plots, echo = FALSE}
##################################
#Tree Plot of Reactome Categories for UP and DOWN gene sets representing Cluster 2 and Cluster 0 respectively
###UP
#0. Define groups of Significant REACTOME categories using a p value cut-off
clean_UP_Sig_pthresh_0.01 <- 
  REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP@result %>%
  filter(p.adjust < 10^-2)

#1. Pairwise clustering of REACTOME categories
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_PW                 <- pairwise_termsim(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP, showCategory = nrow(clean_UP_Sig_pthresh_0.01))

#2. Assign tree.plot and tree.plot_reference
# _UP_tree.plot
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_Tree.pthresh_0.01.plot           <- treeplot(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_PW, showCategory = nrow(clean_UP_Sig_pthresh_0.01))
#_UP_tree.plot_reference for plotting with labels and guides
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_Tree.pthresh_0.01.plot_reference <- treeplot(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_PW, showCategory = nrow(clean_UP_Sig_pthresh_0.01))



#3. modify _UP_tree.plot aesthetics for plotting
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_Tree.pthresh_0.01.plot <- REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_Tree.pthresh_0.01.plot + 
                                                                                                    geom_tree(size=1) +
                                                                                                    geom_tiplab(alpha = 0, size = 0)+
                                                                                                    scale_size_continuous(range = c(1,15))
                                                                                
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_Tree.pthresh_0.01.plot$layers[[7]]$aes_params$size <- 0


#4. Save tree.plot and tree.plot_reference
ggsave(paste0(astro.gsea.path, "REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_Tree.pthresh_0.01.plot.png"), 
       plot = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_Tree.pthresh_0.01.plot, 
       device='png', width = 6, height=12, units='in', dpi=600)

ggsave(paste0(astro.gsea.path, "REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_Tree.pthresh_0.01.plot_reference.png"), 
       plot = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_Tree.pthresh_0.01.plot_reference, 
       device='png', width = 16, height=12, units='in', dpi=600)





#DOWN
#0. Define groups of Significant REACTOME categories using a p value cut-off
clean_DOWN_Sig_pthresh_0.01 <- 
  REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN@result %>%
  filter(p.adjust < 10^-2)

clean_DOWN_Sig_pthresh_0.000001 <- 
  REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN@result %>%
  filter(p.adjust < 10^-6)
                 
#1. Pairwise clustering of REACTOME categories
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_pthresh_0.01.PW     <- pairwise_termsim(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN,  showCategory = nrow(clean_DOWN_Sig_pthresh_0.01))
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_pthresh_0.000001.PW <- pairwise_termsim(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN,  showCategory = nrow(clean_DOWN_Sig_pthresh_0.000001))


#2. Assign tree.plot and tree.plot_reference
# _DOWN_tree.plot
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.01.plot           <- treeplot(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_pthresh_0.01.PW, showCategory = nrow(clean_DOWN_Sig_pthresh_0.01))
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.01.plot_reference <- treeplot(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_pthresh_0.01.PW, showCategory = nrow(clean_DOWN_Sig_pthresh_0.01))

REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.000001.plot           <- treeplot(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_pthresh_0.000001.PW, showCategory = nrow(clean_DOWN_Sig_pthresh_0.000001))
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.000001.plot_reference <- treeplot(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_pthresh_0.000001.PW, showCategory = nrow(clean_DOWN_Sig_pthresh_0.000001))


#3. modify _DOWN_tree.plot aesthetics for plotting
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.01.plot <- REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.01.plot + 
                                                                                                        geom_tree(size=2) +
                                                                                                        geom_tiplab(alpha = 0, size = 0)+
                                                                                                        scale_size_continuous(range = c(5,25))
                                                                                
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.01.plot$layers[[7]]$aes_params$size <- 0


REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.000001.plot <- REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.000001.plot + 
                                                                                                        geom_tree(size=2) +
                                                                                                        geom_tiplab(alpha = 0, size = 0)+
                                                                                                        scale_size_continuous(range = c(5,25))
                                                                                
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.000001.plot$layers[[7]]$aes_params$size <- 0



ggsave(paste0(astro.gsea.path, "REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.01.plot.png"), 
       plot = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.01.plot, 
       device='png', width = 10, height=30, units='in', dpi=600)
ggsave(paste0(astro.gsea.path, "REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.01.plot_reference.png"), 
       plot = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.01.plot_reference, 
       device='png', width = 10, height=30, units='in', dpi=600)

ggsave(paste0(astro.gsea.path, "REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.000001.plot.png"), 
       plot = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.000001.plot, 
       device='png', width = 10, height=50, units='in', dpi=600, limitsize = FALSE)
ggsave(paste0(astro.gsea.path, "REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.000001.plot_reference.png"), 
       plot = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_Tree.pthresh_0.000001.plot_reference, 
       device='png', width = 10, height=50, units='in', dpi=600, limitsize = FALSE)

##################################
```

#4d. REACTOME Heatplots
```{r gene-centric heatplots for specific REACTOME categories, echo=FALSE}
#NOT CURRENTLY USED IN THE MANUSCRIPT 24/9/5
#Heatplots for Genes associated with Reactome categores in the UP and DOWN gene sets
#UP
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP.Full.heatplot   <- heatplot(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP,   foldChange = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_avg_logFC, showCategory=50)
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP.Full.heatplot   <- REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP.Full.heatplot + scale_fill_gradient2(name = "fold change", low = "#006837", mid = "#e6f598", high = "#d73027")

ggsave(paste0(astro.gsea.path, "REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP.Full.heatplot.png"), plot = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP.Full.heatplot, device='png', width = 20, height=15, units='in', dpi=600)

#DOWN
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN.Full.heatplot <- heatplot(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN, foldChange = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_avg_logFC, showCategory=50)
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN.Full.heatplot <- REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN.Full.heatplot + scale_fill_gradient2(name = "fold change", low = "#006837", mid = "#e6f598", high = "#d73027")

ggsave(paste0(astro.gsea.path, "REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN.Full.heatplot.png"), plot = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN.Full.heatplot, device='png', width = 60, height=15, units='in', dpi=600, limitsize = FALSE)


#subset the full REACTOME results for a smaller heat map with representative genes
#UP
KeyDescriptions_UP <- c("Toll-like Receptor Cascades", "tRNA processing",  "rRNA processing in the mitochondrion", "rRNA modification in the nucleus and cytosol", "Transcriptional Regulation by TP53", "Nucleotide-binding domain, leucine rich repeat containing receptor (NLR) signaling pathways")

REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_DescriptionSubset <- REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP

REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_DescriptionSubset@result <- REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP@result[REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP@result$Description %in% c(KeyDescriptions_UP),]


REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_DescriptionSubset.heatplot   <- heatplot(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_DescriptionSubset,   foldChange = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_avg_logFC, showCategory=50)
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_DescriptionSubset.heatplot   <- REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_DescriptionSubset.heatplot + 
                                                                                                            scale_fill_gradient2(name = "fold change", low = "#006837", mid = "#e6f598", high = "#d73027") +
                                                                                                            coord_flip()+
                                                                                                            scale_x_discrete(limits=rev)+
                                                                                                            #scale_x_reverse()+
                                                                                                            theme_dark()+
                                                                                                            guides(fill = "none")+
                                                                                                              theme(axis.text.x = element_text(face = "italic", size = 9, angle = 90, family = "Arial"),
                                                                                                                    axis.text.y = element_text(face = "italic", size = 8.5, family = "Arial"))

ggsave(paste0(astro.gsea.path, "REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_DescriptionSubset.heatplot.png"), plot = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_UP_DescriptionSubset.heatplot, device='png', width = 2, height=16, units='in', dpi=600)


#DOWN
KeyDescriptions_DOWN <- c("SARS-CoV Infections", "Viral mRNA Translation", "HIV Infection", "Translation", "Signaling by ROBO receptors", "The citric acid (TCA) cycle and respiratory electron transport", "Metabolism of amino acids and derivatives", "Metabolism of carbohydrates")

REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_DescriptionSubset <- REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN

REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_DescriptionSubset@result <- REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN@result[REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN@result$Description %in% c(KeyDescriptions_DOWN),]


REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_DescriptionSubset.heatplot   <- heatplot(REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_DescriptionSubset,   foldChange = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_avg_logFC, showCategory=50)
REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_DescriptionSubset.heatplot   <- REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_DescriptionSubset.heatplot + 
                                                                                                            scale_fill_gradient2(name = "fold change", low = "#006837", mid = "#e6f598", high = "#d73027") +
                                                                                                            coord_flip()+
                                                                                                            scale_x_discrete(limits=rev)+
                                                                                                            #scale_x_reverse()+
                                                                                                            theme_dark()+
                                                                                                            guides(fill = "none")+
                                                                                                              theme(axis.text.x = element_text(face = "italic", size = 10, angle = 90, family = "Arial"),
                                                                                                                    axis.text.y = element_text(face = "italic", size = 10, family = "Arial"))

ggsave(paste0(astro.gsea.path, "REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_DescriptionSubset.heatplot.png"), plot = REACTOME.Pathway.DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_clean_DOWN_DescriptionSubset.heatplot, device='png', width = 3, height=80, units='in', dpi=600, limitsize = FALSE)


```

#4e. Custom REACTOME Heatplots for whole Gene Categories
```{r custom heatplots, echo = FALSE}

#0a. Load REACTOME Gene Lists for Key Categories
RHSA168164_TLR                 <- read_tsv(REACTOME_RHSA168164_TLR.path)
RHSA72766_Translation          <- read_tsv(REACTOME_RHSA72766_Translation.path)
RHSA3700989_TP53               <- read_tsv(REACTOME_RHSA3700989_TP53.path)
RHSA72306_tRNA                 <- read_tsv(REACTOME_RHSA72306_tRNA.path)
RHSA611105_REC                 <- read_tsv(REACTOME_RHSA611105_REC.path)

#0b.Extract Gene Names
RHSA168164_TLR_Genes        <- sapply(strsplit(RHSA168164_TLR$MoleculeName,split=" "),       tail,1)
RHSA72766_Translation_Genes <- sapply(strsplit(RHSA72766_Translation$MoleculeName,split=" "),tail,1)
RHSA3700989_TP53_Genes      <- sapply(strsplit(RHSA3700989_TP53$MoleculeName,split=" "),     tail,1)
RHSA72306_tRNA_Genes        <- sapply(strsplit(RHSA72306_tRNA$MoleculeName,split=" "),       tail,1)
RHSA611105_REC_Genes        <- sapply(strsplit(RHSA611105_REC$MoleculeName,split=" "),       tail,1)
####


#1. Toll Like Receptor (9RHSA168164)
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA168164_TLR_GeneSet <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All %>%
  filter(SYMBOL %in% RHSA168164_TLR_Genes)  %>%
  filter(!is.na(avg_logFC)) %>%
  dplyr::select(SYMBOL, avg_logFC, p_val_adj) %>%
  arrange(desc(avg_logFC)) %>%
  mutate(avg_logFC_tile = ntile(avg_logFC, 4)) %>%
  mutate(IsSig = case_when(p_val_adj < 0.01 ~ "yes", TRUE ~ "no")) %>%
  mutate(IsSig = factor(IsSig, levels = c("yes", "no"))) %>%
  mutate(SYMBOL = factor(SYMBOL, 
                     ordered=TRUE, 
                     levels=unique(.[order(.$avg_logFC),"SYMBOL"][[1]])))

HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA168164_TLR_Genes <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA168164_TLR_GeneSet, aes(y = SYMBOL, x = 1, fill = avg_logFC))+
    geom_tile(color = "black")+
    geom_point(aes(size = IsSig, alpha = IsSig))+
    coord_equal()+
    scale_fill_gradient2(limits = c(-3,3), low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0, space = "Lab", na.value = muted("red"))+
    scale_size_discrete(range = c(1, 4))+
    scale_alpha_discrete(range = c(1, 0))+
    theme(axis.text.y = element_text(family = "Arial", size = 12, face = "italic"))

ggsave(paste0(astro.gsea.path, "HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA168164_TLR_Genes.png"), 
       plot = HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA168164_TLR_Genes, 
       device='png', width = 3, height=16, units='in', dpi=600, limitsize = FALSE)
####

####
#2. Translation (RHSA72766)
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All %>%
  filter(SYMBOL %in% RHSA72766_Translation_Genes)  %>%
  filter(!is.na(avg_logFC)) %>%
  dplyr::select(SYMBOL, avg_logFC, p_val_adj) %>%
  arrange(desc(avg_logFC)) %>%
  mutate(avg_logFC_tile = ntile(avg_logFC, 4)) %>%
  mutate(IsSig = case_when(p_val_adj < 0.01 ~ "yes", TRUE ~ "no")) %>%
  mutate(IsSig = factor(IsSig, levels = c("yes", "no"))) %>%
  mutate(SYMBOL = factor(SYMBOL, 
                     ordered=TRUE, 
                     levels=unique(.[order(.$avg_logFC),"SYMBOL"][[1]])))

HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA72766_Translation_Genes <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet, aes(y = SYMBOL, SYMBOL, x = 1, fill = avg_logFC))+
    geom_tile(color = "black")+
    geom_point(aes(size = IsSig, alpha = IsSig))+
    coord_equal()+
    scale_fill_gradient2(limits = c(-3,3), low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0, space = "Lab", na.value = muted("red"))+
    scale_size_discrete(range = c(1, 4))+
    scale_alpha_discrete(range = c(1, 0))+
    theme(axis.text.y = element_text(family = "Arial", size = 12, face = "italic"))

ggsave(paste0(astro.gsea.path, "HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA72766_Translation_Genes.png"), 
       plot = HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA72766_Translation_Genes, 
       device='png', width = 3, height=40, units='in', dpi=600, limitsize = FALSE)
####

####
#3. TP53 (RHSA3700989)
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA3700989_TP53_GeneSet <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All %>%
  filter(SYMBOL %in% RHSA3700989_TP53_Genes)  %>%
  filter(!is.na(avg_logFC)) %>%
  dplyr::select(SYMBOL, avg_logFC, p_val_adj) %>%
  arrange(desc(avg_logFC)) %>%
  mutate(avg_logFC_tile = ntile(avg_logFC, 4)) %>%
  mutate(IsSig = case_when(p_val_adj < 0.01 ~ "yes", TRUE ~ "no")) %>%
  mutate(IsSig = factor(IsSig, levels = c("yes", "no"))) %>%
  mutate(SYMBOL = factor(SYMBOL, 
                     ordered=TRUE, 
                     levels=unique(.[order(.$avg_logFC),"SYMBOL"][[1]])))

HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA3700989_TP53_Genes <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA3700989_TP53_GeneSet, aes(y = SYMBOL, SYMBOL, x = 1, fill = avg_logFC))+
    geom_tile(color = "black")+
    geom_point(aes(size = IsSig, alpha = IsSig))+
    coord_equal()+
    scale_fill_gradient2(limits = c(-3,3), low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0, space = "Lab", na.value = muted("red"))+
    scale_size_discrete(range = c(1, 4))+
    scale_alpha_discrete(range = c(1, 0))+
    theme(axis.text.y = element_text(family = "Arial", size = 18, face = "italic"))

ggsave(paste0(astro.gsea.path, "HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA3700989_TP53_Genes.png"), 
       plot = HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA3700989_TP53_Genes, 
       device='png', width = 3, height=40, units='in', dpi=600, limitsize = FALSE)
####

####
#5. tRNA (RHSA72306)
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72306_tRNA_GeneSet <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All %>%
  filter(SYMBOL %in% RHSA72306_tRNA_Genes)  %>%
  filter(!is.na(avg_logFC)) %>%
  dplyr::select(SYMBOL, avg_logFC, p_val_adj) %>%
  arrange(desc(avg_logFC)) %>%
  mutate(avg_logFC_tile = ntile(avg_logFC, 4)) %>%
  mutate(IsSig = case_when(p_val_adj < 0.01 ~ "yes", TRUE ~ "no")) %>%
  mutate(IsSig = factor(IsSig, levels = c("yes", "no"))) %>%
  mutate(SYMBOL = factor(SYMBOL, 
                     ordered=TRUE, 
                     levels=unique(.[order(.$avg_logFC),"SYMBOL"][[1]])))

HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA72306_tRNA_Genes <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72306_tRNA_GeneSet, aes(y = SYMBOL, SYMBOL, x = 1, fill = avg_logFC))+
    geom_tile(color = "black")+
    geom_point(aes(size = IsSig, alpha = IsSig))+
    coord_equal()+
    scale_fill_gradient2(limits = c(-3,3), low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0, space = "Lab", na.value = muted("red"))+
    scale_size_discrete(range = c(1, 4))+
    scale_alpha_discrete(range = c(1, 0))+
    theme(axis.text.y = element_text(family = "Arial", size = 18, face = "italic"))

ggsave(paste0(astro.gsea.path, "HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA72306_tRNA_Genes.png"), 
       plot = HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA72306_tRNA_Genes, 
       device='png', width = 3, height=40, units='in', dpi=600, limitsize = FALSE)
####


####
#6. REC (RHSA611105)
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA611105_REC_GeneSet <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All %>%
  filter(SYMBOL %in% RHSA611105_REC_Genes)  %>%
  filter(!is.na(avg_logFC)) %>%
  dplyr::select(SYMBOL, avg_logFC, p_val_adj) %>%
  arrange(desc(avg_logFC)) %>%
  mutate(avg_logFC_tile = ntile(avg_logFC, 4)) %>%
  mutate(IsSig = case_when(p_val_adj < 0.01 ~ "yes", TRUE ~ "no")) %>%
  mutate(IsSig = factor(IsSig, levels = c("yes", "no"))) %>%
  mutate(SYMBOL = factor(SYMBOL, 
                     ordered=TRUE, 
                     levels=unique(.[order(.$avg_logFC),"SYMBOL"][[1]])))

HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA611105_REC_Genes <-
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA611105_REC_GeneSet, aes(y = SYMBOL, SYMBOL, x = 1, fill = avg_logFC))+
    geom_tile(color = "black")+
    geom_point(aes(size = IsSig, alpha = IsSig))+
    coord_equal()+
    scale_fill_gradient2(limits = c(-3,3), low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0, space = "Lab", na.value = muted("red"))+
    scale_size_discrete(range = c(1, 4))+
    scale_alpha_discrete(range = c(1, 0))+
    theme(axis.text.y = element_text(family = "Arial", size = 18, face = "italic"))

ggsave(paste0(astro.gsea.path, "HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA611105_REC_Genes.png"), 
       plot = HeatMapLog2FC_AstroLow_Cluster0.vs.AstroHigh_Cluster2_RHSA611105_REC_Genes, 
       device='png', width = 3, height=40, units='in', dpi=600, limitsize = FALSE)
####

```

```{r preprocess DE.tha05.vs.NI.CellType for comparison to , echo=FALSE}

#filter to just Astrocytes
DE.tha05.vs.NI.CellType.Astro <-
  DE.tha05.vs.NI.CellType %>%
  filter(cell_type == "Astrocyte")


```

#REACTOME Category inspired Gene Set Analysis
```{r focused analysis on 3 gene sets: ribosome, tRNA, cox_nduf, echo = F}

#Ribosome
mito_ribo    <- grep(pattern = "^MRP*", x = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet$SYMBOL, perl = T, value = TRUE)
nuclear_ribo <- unique(x = grep(pattern = "^RPS*", x = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet$SYMBOL, perl = T, value = TRUE), y = grep(pattern = "^RPL*", x = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet$SYMBOL, perl = T, value = TRUE))

#tRNA
mito_tRNA    <- c("MARS2", "EARS2", "FARS2", "HARS2", "PARS2", "SARS2", "LARS2", "YARS2", "VARS2", "WARS2", "AARS2", "NARS2", "RARS2", "CARS2", "TARS2", "DARS2", "IARS2")
nuclear_tRNA <- c("AARS", "CARS", "DARS", "EPRS", "FARSA", "FARSB", "IARS",  "LARS", "MARS", "NARS", "QARS", "RARS", "SARS", "TARS", "VARS", "WARS", "YARS")
both_tRNA    <- c("GARS", "KARS", "HARS")

#Electron Transport
cox     <- grep(pattern = "^COX", x = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All$SYMBOL, perl = T, value = TRUE)
nduf    <- grep(pattern = "^NDUF", x = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All$SYMBOL, perl = T, value = TRUE)

#DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_Ribo <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta1 %>%
    mutate(RiboGeneCategory = case_when(gene %in% mito_ribo ~ "Mitochondria", gene %in% nuclear_ribo ~ "Cytoplasmic", TRUE ~ NA)) %>%
    filter(!is.na(RiboGeneCategory))%>%
    mutate(IsSig = case_when(p_val_adj < 0.01 ~ "yes", TRUE ~ "no")) %>%
    mutate(IsSig = factor(IsSig, levels = c("yes", "no")))

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_tRNA <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta1 %>%
    mutate(tRNAGeneCategory = case_when(gene %in% mito_tRNA ~ "Mitochondria", gene %in% nuclear_tRNA ~ "Cytoplasmic", gene %in% both_tRNA ~ "Both", TRUE ~ NA)) %>%
    filter(!is.na(tRNAGeneCategory)) %>%
    mutate(IsSig = case_when(p_val_adj < 0.01 ~ "yes", TRUE ~ "no")) %>%
    mutate(IsSig = factor(IsSig, levels = c("yes", "no")))

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_COX_NDUF_GeneSet_REC <-
  DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_meta1 %>%
    mutate(COX_NDUF_GeneCategory = case_when(gene %in% cox ~ "COX", gene %in% nduf ~ "NDUF", TRUE ~ NA)) %>%
    filter(!is.na(COX_NDUF_GeneCategory)) %>%
    mutate(IsSig = case_when(p_val_adj < 0.01 ~ "yes", TRUE ~ "no")) %>%
    mutate(IsSig = factor(IsSig, levels = c("yes", "no"))) %>%
    filter(!grepl("-AS",gene)) %>%
    filter(!grepl("-DT",gene))

###Plotting
#######################DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2###################################
#dot parameters
dot_size = 5
dot_stroke = 1.2
dot_alpha = 1


#Ribosomal Genes


DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_Ribo.jitter_reference.plot <- 
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_Ribo, aes(x=RiboGeneCategory, y=avg_logFC, color = IsSig, fill = RiboGeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha)+               #geom_point(position = position_jitter(seed = 42))
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("red", "black"))+
    scale_fill_manual(values = c("#ccece6", "#66c2a4"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()

ggsave(paste0(astro.gsea.path, "DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_Ribo.jitter_reference.plot.png"), 
       plot = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_Ribo.jitter_reference.plot, 
       device='png', width = 5, height=5, units='in', dpi=600, limitsize = FALSE)


DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_Ribo.jitter.plot <- 
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_Ribo, aes(x=RiboGeneCategory, y=avg_logFC, color = IsSig, fill = RiboGeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha, stroke = dot_stroke)+
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("red", "black"))+
    scale_fill_manual(values = c("#ccece6", "#66c2a4"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()+
    guides(color = "none", fill = "none")

ggsave(paste0(astro.gsea.path, "DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_Ribo.jitter.plot.png"), 
       plot = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_Ribo.jitter.plot, 
       device='png', width = 3.5, height=5, units='in', dpi=600, limitsize = FALSE)

#tRNAs Genes
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_tRNA.jitter_reference.plot <- 
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_tRNA, aes(x=tRNAGeneCategory, y=avg_logFC, color = IsSig, fill = tRNAGeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha)+
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("red", "black"))+
    scale_fill_manual(values = c("#ffffe5", "#fff7bc", "#fee391"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()

ggsave(paste0(astro.gsea.path, "DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_tRNA.jitter_reference.plot.png"), 
       plot = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_tRNA.jitter_reference.plot, 
       device='png', width = 5, height=5, units='in', dpi=600, limitsize = FALSE)


DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_tRNA.jitter.plot <- 
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_tRNA, aes(x=tRNAGeneCategory, y=avg_logFC, color = IsSig, fill = tRNAGeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha)+
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("red", "black"))+
    scale_fill_manual(values = c("#ffffe5", "#fff7bc", "#fee391"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()+
    guides(color = "none", fill = "none")

ggsave(paste0(astro.gsea.path, "DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_tRNA.jitter.plot.png"), 
       plot = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_RHSA72766_Translation_GeneSet_tRNA.jitter.plot, 
       device='png', width = 3.5, height=5, units='in', dpi=600, limitsize = FALSE)


#COX_NDUF Genes
DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_COX_NDUF_GeneSet_REC.jitter_reference.plot <- 
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_COX_NDUF_GeneSet_REC, aes(x=COX_NDUF_GeneCategory, y=avg_logFC, color = IsSig, fill = COX_NDUF_GeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha)+
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("red", "black"))+
    scale_fill_manual(values = c("#deebf7", "#9ecae1"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()

ggsave(paste0(astro.gsea.path, "DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_COX_NDUF_GeneSet_REC.jitter_reference.plot.png"), 
       plot = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_COX_NDUF_GeneSet_REC.jitter_reference.plot, 
       device='png', width = 5, height=5, units='in', dpi=600, limitsize = FALSE)

DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_COX_NDUF_GeneSet_REC.jitter.plot <- 
  ggplot(DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_COX_NDUF_GeneSet_REC, aes(x=COX_NDUF_GeneCategory, y=avg_logFC, color = IsSig, fill = COX_NDUF_GeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha)+
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("red", "black"))+
    scale_fill_manual(values = c("#deebf7", "#9ecae1"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()+
    guides(color = "none", fill = "none")

ggsave(paste0(astro.gsea.path, "DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_COX_NDUF_GeneSet_REC.jitter.plot.png"), 
       plot = DE.AstroLow_Cluster0.vs.AstroHigh_Cluster2_All_COX_NDUF_GeneSet_REC.jitter.plot, 
       device='png', width = 3.5, height=5, units='in', dpi=600, limitsize = FALSE)

##################################################################################

#######################DE.tha05.vs.NI.CellType.Astro##############################

#DE.tha05.vs.NI.CellType.Astro
DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_Ribo <-
  DE.tha05.vs.NI.CellType.Astro %>%
    mutate(RiboGeneCategory = case_when(gene %in% mito_ribo ~ "Mitochondria", gene %in% nuclear_ribo ~ "Cytoplasmic", TRUE ~ NA)) %>%
    filter(!is.na(RiboGeneCategory))%>%
    mutate(IsSig = case_when(p_val_adj < 0.01 ~ "yes", TRUE ~ "no")) %>%
    mutate(IsSig = factor(IsSig, levels = c("yes", "no")))

DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_tRNA <-
  DE.tha05.vs.NI.CellType.Astro %>%
    mutate(tRNAGeneCategory = case_when(gene %in% mito_tRNA ~ "Mitochondria", gene %in% nuclear_tRNA ~ "Cytoplasmic", gene %in% both_tRNA ~ "Both", TRUE ~ NA)) %>%
    filter(!is.na(tRNAGeneCategory)) %>%
    mutate(IsSig = case_when(p_val_adj < 0.01 ~ "yes", TRUE ~ "no")) %>%
    mutate(IsSig = factor(IsSig, levels = c("yes", "no")))

DE.tha05.vs.NI.CellType.Astro_COX_NDUF_GeneSet_REC <-
  DE.tha05.vs.NI.CellType.Astro %>%
    mutate(COX_NDUF_GeneCategory = case_when(gene %in% cox ~ "COX", gene %in% nduf ~ "NDUF", TRUE ~ NA)) %>%
    filter(!is.na(COX_NDUF_GeneCategory)) %>%
    mutate(IsSig = case_when(p_val_adj < 0.01 ~ "yes", TRUE ~ "no")) %>%
    mutate(IsSig = factor(IsSig, levels = c("yes", "no"))) %>%
    filter(!grepl("-AS",gene)) %>%
    filter(!grepl("-DT",gene))





###Plotting
#Ribosomal Genes

DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_Ribo.jitter_reference.plot <- 
  ggplot(DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_Ribo, aes(x=RiboGeneCategory, y=avg_logFC, color = IsSig, fill = RiboGeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha)+               #geom_point(position = position_jitter(seed = 42))
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("red", "black"))+
    scale_fill_manual(values = c("#ccece6", "#66c2a4"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()

ggsave(paste0(astro.gsea.path, "DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_Ribo.jitter_reference.plot.png"), 
       plot = DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_Ribo.jitter_reference.plot, 
       device='png', width = 5, height=5, units='in', dpi=600, limitsize = FALSE)


DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_Ribo.jitter.plot <- 
  ggplot(DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_Ribo, aes(x=RiboGeneCategory, y=avg_logFC, color = IsSig, fill = RiboGeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha)+
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("red", "black"))+
    scale_fill_manual(values = c("#ccece6", "#66c2a4"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()+
    guides(color = "none", fill = "none")

ggsave(paste0(astro.gsea.path, "DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_Ribo.jitter.plot.png"), 
       plot = DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_Ribo.jitter.plot, 
       device='png', width = 3.5, height=5, units='in', dpi=600, limitsize = FALSE)

#tRNAs Genes
DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_tRNA.jitter_reference.plot <- 
  ggplot(DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_tRNA, aes(x=tRNAGeneCategory, y=avg_logFC, color = IsSig, fill = tRNAGeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha)+
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("red", "black"))+
    scale_fill_manual(values = c("#ffffe5", "#fff7bc", "#fee391"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()

ggsave(paste0(astro.gsea.path, "DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_tRNA.jitter_reference.plot.png"), 
       plot = DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_tRNA.jitter_reference.plot, 
       device='png', width = 5, height=5, units='in', dpi=600, limitsize = FALSE)


DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_tRNA.jitter.plot <- 
  ggplot(DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_tRNA, aes(x=tRNAGeneCategory, y=avg_logFC, color = IsSig, fill = tRNAGeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha)+
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("black", "red"))+
    scale_fill_manual(values = c("#ffffe5", "#fff7bc", "#fee391"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()+
    guides(color = "none", fill = "none")

ggsave(paste0(astro.gsea.path, "DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_tRNA.jitter.plot.png"), 
       plot = DE.tha05.vs.NI.CellType.Astro_RHSA72766_Translation_GeneSet_tRNA.jitter.plot, 
       device='png', width = 3.5, height=5, units='in', dpi=600, limitsize = FALSE)


#COX_NDUF Genes
DE.tha05.vs.NI.CellType.Astro_COX_NDUF_GeneSet_REC.jitter_reference.plot <- 
  ggplot(DE.tha05.vs.NI.CellType.Astro_COX_NDUF_GeneSet_REC, aes(x=COX_NDUF_GeneCategory, y=avg_logFC, color = IsSig, fill = COX_NDUF_GeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha)+
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("black", "red"))+
    scale_fill_manual(values = c("#deebf7", "#9ecae1"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()

ggsave(paste0(astro.gsea.path, "DE.tha05.vs.NI.CellType.Astro_COX_NDUF_GeneSet_REC.jitter_reference.plot.png"), 
       plot = DE.tha05.vs.NI.CellType.Astro_COX_NDUF_GeneSet_REC.jitter_reference.plot, 
       device='png', width = 5, height=5, units='in', dpi=600, limitsize = FALSE)

DE.tha05.vs.NI.CellType.Astro_COX_NDUF_GeneSet_REC.jitter.plot <- 
  ggplot(DE.tha05.vs.NI.CellType.Astro_COX_NDUF_GeneSet_REC, aes(x=COX_NDUF_GeneCategory, y=avg_logFC, color = IsSig, fill = COX_NDUF_GeneCategory))+
    geom_violin(color = "black", fill = "#f0f0f0")+
    geom_jitter(shape = 21, size = dot_size, stroke = dot_stroke, alpha = dot_alpha)+
    scale_y_continuous(limits = c(-2,2))+
    scale_color_manual(values = c("black", "red"))+
    scale_fill_manual(values = c("#deebf7", "#9ecae1"))+
    geom_hline(yintercept = 0, linewidth = 1, linetype = "dotted")+
    theme_bw()+
    guides(color = "none", fill = "none")

ggsave(paste0(astro.gsea.path, "DE.tha05.vs.NI.CellType.Astro_COX_NDUF_GeneSet_REC.jitter.plot.png"), 
       plot = DE.tha05.vs.NI.CellType.Astro_COX_NDUF_GeneSet_REC.jitter.plot, 
       device='png', width = 3.5, height=5, units='in', dpi=600, limitsize = FALSE)





##################################################################################
```
